---
title: "p8105_hw2_gw2442"
output: github_document
author: "Gwyneth Wei"
---
```{r}
library(tidyverse)
library(readxl)
```

## Problem 1 

Reading and cleaning data in `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`:

```{r}
transit_df = read_csv(
  'data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv',
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line:vending, ada) %>%
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE))

skimr::skim(transit_df)
```

The dataset `transit_df` contains variables line, station_name, route1 - route11, entrance_type, vending, entry, exit_only, ada, station_latitude, and station_longitude. 

Following the import of the csv data, a new dataset was created called `transit_df` where the column names were cleaned using `janitor::clean_names()`. The variables route8 - route11 were converted to character variables. The variables of interest were selected from the original csv file for the new dataset and the variable `entry` was converted to a logical variable using `ifelse`. The dimensions of the resulting dataset are 1868 rows and 20 columns. The data is not tidy, as the columns route1 - route11 are values rather than variable names. 

*How many distinct stations are there?*
```{r}
select(transit_df, station_name, line) %>%
distinct
```
There are 465 distinct stations.

*How many stations are ADA compliant?*
```{r}
transit_df %>%
    filter(ada == TRUE) %>%
    select(station_name, line) %>%
    distinct
```
There are 84 stations that are ADA compliant.

*What proportion of station entrances/exits without vending allow entrance?*
```{r}
transit_df %>%
    filter(vending == "NO") %>%
    pull(entry) %>%
    mean
```
37.70% of all station entrances/exits without vending allow entrance.

*How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?*
```{r}
transit_df %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name") %>%
  filter(route_name == "A") %>%
  select (station_name, line) %>%
  distinct

transit_df %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name") %>%
  filter(route_name == "A", ada == TRUE) %>%
  select(station_name, line) %>%
  distinct
```

There are 60 distinct stations that serve the A train. Of the stations that serve the A train, there are 17 that are ADA compliant. 


## Problem 2

Reading and cleaning Mr. Trash Wheel sheet:

* Variable `trash_wheel` was added for combining the datasets later on
```{r}
m_trash_wheel = readxl::read_excel(
  "data/Trash Wheel Collection Data.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N549") %>%
  janitor::clean_names() %>%
  mutate(
    sports_balls = round(sports_balls),
    trash_wheel = "mr") %>%
  select(trash_wheel, everything())
```

Reading and cleaning Professor Trash Wheel sheet:

* Variable `trash_wheel` was added for combining the datasets later on
* Variable `year` was converted into a character variable in order to combine the datasets later on 
```{r}
p_trash_wheel = readxl::read_excel(
  "data/Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel",
  range = "A2:M96") %>%
  janitor::clean_names() %>%
  mutate(
    year = as.character(year),
    trash_wheel = "professor") %>%
  select(trash_wheel, everything())
```

Combining datasets `m_trash_wheel` and `p_trash_wheel`
```{r}
trash_wheel_tidy = bind_rows(m_trash_wheel, p_trash_wheel)

skimr::skim(trash_wheel_tidy)
```
The dataset `trash_wheel_tidy` has 641 observations (547 from dataset `m_trash_wheel` and 94 from `p_trash_wheel`). There are 15 variables, including the variable `trash_wheel` that was added to keep track of which Trash Wheel is which (Mr. or Professor). Some of the key variables in this dataset include `dumpster`, `weight_tons`, `volume_cubic_yards`, and `sports_balls`. 

*What was the total weight of trash collected by Professor Trash Wheel? What was the total number of sports balls collected by Mr. Trash Wheel in 2020?*
```{r}
aggregate (x = trash_wheel_tidy$weight_tons,
           by = list(trash_wheel_tidy$trash_wheel),
           FUN = sum)

aggregate (x = trash_wheel_tidy$sports_balls,
           by = list(trash_wheel_tidy$trash_wheel, trash_wheel_tidy$year),
           FUN = sum)
```

The total weight of trash collected by Professor Trash Wheel is 190.12  tons.
The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856 sports balls. 



## Problem 3 
Cleaning data in `pols-month.csv`
```{r}
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c('year', 'month', 'day'), sep = "-") %>%
  mutate( 
    month = recode(month,
    "01" = "jan", "02" = "feb", "03" = "mar", "04" = "apr", "05" = "may", "06" = "jun", "07" = "jul", "08" = "aug", "09" = "sep", "10" = "oct", "11" = "nov", "12" = "dec"),
    prez_dem = recode(prez_dem, "1" = "dem"),
    prez_gop = recode(prez_gop, "1" = "gop")) %>%
   pivot_longer(
     cols = starts_with("prez"),
     names_to = "temp",
     values_to = "president") %>%
  drop_na("president") %>%
  select(-temp, -day)
```

Cleaning data in `snp.csv`
```{r}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(
    date = as.Date(date, "%m/%d/%y"),
    date = as.Date(ifelse(date > "2015-07-01", format(date, "19%y-%m-%d"), format(date)))) %>%
  separate(date, c('year', 'month', 'day'), sep = "-") %>%
  select(year, month, day, everything()) %>%
  arrange(year, month) %>%
  select(-day) %>%
  mutate(
    month = recode(month,
    "01" = "jan", "02" = "feb", "03" = "mar", "04" = "apr", "05" = "may", "06" = "jun", "07" = "jul", "08" = "aug", "09" = "sep", "10" = "oct", "11" = "nov", "12" = "dec"))
```

Cleaning data in `unemployment.csv` 

* Variable `year` was converted to a character variable for combining datasets later on
```{r}
unemployment_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "pct_unemployment") 

unemployment_df$year <- as.character(unemployment_df$year)
```

Joining `snp_df` into `pols_df`, and merging`unemployment_df` with the result
```{r}
pols_snp_combined =
  left_join(pols_df, snp_df, by = c("year", "month"))
view(pols_snp_combined)

problem3_tidy = 
  left_join(pols_snp_combined, unemployment_df, by = c("year", "month"))
view(problem3_tidy)

skimr::skim(problem3_tidy)
```

After cleaning the `pols-month.csv`, the `pols_df` dataset contained 817 rows and 9 columns. The variables within the dataset included year, month, president, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, and rep_dem. The variable year ranged from 1947-2015.

After cleaning the `snp.csv`, the `snp_df` dataset 
contained 787 rows and 3 columns. The variables within the dataset included year, month, and close. The variable year ranged from 
1950-2015.

After cleaning the `unemployment.csv`, the `unemployment_df` dataset contained 816 rows and 3 columns. The variables within the dataset included year, month, and pct_unemployment. The variable year ranged from 1948-1950. 

After joining the three datasets into `problem3_tidy`, the resulting datset contained 817 rows and 11 columns. The variables within the dataset included year, month, president, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, close, and pct_unemployment. The variable year ranged from 1947-2015. 






