---
title: "p8105_hw2_gw2442"
output: github_document
date: "2022-09-28"
---
```{r}
library(tidyverse)
```

## Problem 1 

```{r}
transit_df = read_csv('data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE))
    
skimr::skim(transit_df)
```

The dataset contains variables line, station_name, route1 - route11, entrance_type, vending, entry, ada, station_latitude, and station_longitude. Following the import of the csv data, a new dataset was created called `transit_df` where the column names were cleaned using `janitor::clean_names()`. The variables of interest were selected from the original csv file for the new dataset and the variable `entry` was converted to a logical variable using `ifelse`. The dimensions of the resulting dataset are 1868 rows and 19 columns. The data is not tidy, as the columns route1 - route11 are values rather than variable names. 


```{r}
distinct(transit_df, line, station_name)
```
There are 465 distinct stations.

```{r}
skimr::skim(transit_df)
```
There are 468 stations that are ADA compliant.

```{r}
distinct(transit_df, line, station_name, vending)

```
There are 465 station entrances/exists without vending that allow entrance. ***

```{r}
transit_df = 
pivot_longer(
    transit_df, 
    route1:route11,
    values_transform = as.character,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    ) 
view(transit_df)

distinct(transit_df, station_name,line) %>%
    count(vending = "YES", entry = TRUE)
```


## Problem 2

Reading and cleaning Mr. Trash Wheel sheet:
```{r}
m_trash_wheel = readxl::read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = "A2:N533") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    trash_wheel = "mr") %>%
  select(trash_wheel, everything())
```

Reading and cleaning Professor Trash Wheel sheet:
```{r}
p_trash_wheel = readxl::read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Professor Trash Wheel",
  range = "A2:N117") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    trash_wheel = "professor") %>%
  select(trash_wheel, everything())
```

Combining datasets `m_trash_wheel` and `p_trash_wheel`
```{r}
trash_wheel_tidy = bind_rows(m_trash_wheel, p_trash_wheel)

skimr::skim(trash_wheel_tidy)
```
The dataset `trash_wheel_tidy` has 524 observations (453 from dataset `m_trash_tidy` and 71 from `p_trash_tidy`). There are 15 variables, including the variable `trash_wheel` that was added to keep track of which Trash Wheel (Mr. or Professor) is which. Some of the key variables in this dataset include `dumpster`, `weight_tons`, `volume_cubic_yards`, and `sports_balls`. 

```{r}
aggregate (x = trash_wheel_tidy$weight_tons,
           by = list(trash_wheel_tidy$trash_wheel),
           FUN = sum)

aggregate (x = trash_wheel_tidy$sports_balls,
           by = list(trash_wheel_tidy$trash_wheel, trash_wheel_tidy$year),
           FUN = sum)
```

The total weight of trash collected by Professor Trash Wheel is 135.5 tons.
The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856 sports balls. 












